---
title: "Data 624 Homework 8"
subtitle: "Week 11 Non-Linear Regression"
author: "Stephen Haslett"
date: "11/08/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(caret)
library(tidyr)
library(dplyr)
library(mice)
library(kableExtra)
library(mlbench)
library(DT)
library(tsibble)
library(lubridate)
library(ggplot2)
library(AppliedPredictiveModeling)
library(earth)
library(kernlab)
```

### Exercise 7.2

Friedman (1991) introduced several benchmark data sets create by simulation. One of these simulations used the following nonlinear equation to create data:

$y = 10 \sin(\pi x_{1}x_{2}) + 20(x_{3} - 0.5)^2 + 10x_{4} + 5x_{5} + N(0, \sigma^2)$

where the $x$ values are random variables uniformly distributed between [0, 1] (there are also 5 other non-informative variables also created in the simulation). The package mlbench contains a function called _mlbench.friedman1_ that simulates these data:

```{r 7.2Example, eval=TRUE, message=FALSE, warning=FALSE}
set.seed(200)
trainingData <- mlbench.friedman1(200, sd = 1)
## We convert the 'x' data from a matrix to a data frame
## One reason is that this will give the columns names.
trainingData$x <- data.frame(trainingData$x)
## Look at the data using
featurePlot(trainingData$x, trainingData$y)
## or other methods.

## This creates a list with a vector 'y' and a matrix
## of predictors 'x'. Also simulate a large test set to
## estimate the true error rate with good precision:
testData <- mlbench.friedman1(5000, sd = 1)
testData$x <- data.frame(testData$x)
```


#### (a) Tune several models on these data. For example:

```{r 7.2AExample1, eval=TRUE, message=FALSE, warning=FALSE}
knnModel <- train(x = trainingData$x,
                  y = trainingData$y,
                  method = "knn",
                  preProc = c("center", "scale"),
                  tuneLength = 10)
knnModel
```


```{r 7.2AExample2, eval=TRUE, message=FALSE, warning=FALSE}
knnPred <- predict(knnModel, newdata = testData$x)
## The function 'postResample' can be used to get the test set
## performance values.
knnPerformance <- postResample(pred = knnPred, obs = testData$y)
knnPerformance
```


#### Which models appear to give the best performance?

**Model 1: MARS Model**

```{r 7.2MARSModel, eval=TRUE, message=FALSE, warning=FALSE}
set.seed(50)
# Define and tune the MARS model.
marsGrid <- expand.grid(.degree = 1:2, .nprune = 2:20)
marsModel <- train(x = trainingData$x,
                   y = trainingData$y,
                   method = 'earth',
                   tuneGrid = marsGrid,
                   preProc = c('center', 'scale'))

marsModel
```



```{r 7.2MARSModelPredict, eval=TRUE, message=FALSE, warning=FALSE}
# Run predict() and postResample() on the model.
marsPred <- predict(marsModel, newdata = testData$x)
marsPerformance <- postResample(pred = marsPred, obs = testData$y)
marsPerformance
```


**Model 2: SVM Model**

```{r 7.2SVMModel, eval=TRUE, message=FALSE, warning=FALSE}
set.seed(50)
# Define and tune the SVM model.
svmModel <- train(x = trainingData$x,
                  y = trainingData$y,
                  method = 'svmRadial',
                  preProc = c('center', 'scale'),
                  tuneLength = 14,
                  trControl = trainControl(method = 'cv'))

svmModel
```


```{r 7.2SVMModelPredict, eval=TRUE, message=FALSE, warning=FALSE}
# Run predict() and postResample().
svmPred <- predict(svmModel, newdata = testData$x)
svmPerformance <- postResample(pred = svmPred, obs = testData$y)
svmPerformance
```


**Model 3: Neural Network Model**

```{r 7.2NNetModel, eval=TRUE, message=FALSE, warning=FALSE}
set.seed(50)
# Define and tune the Neural Network model.
nnetGrid <- expand.grid(.decay = c(0, 0.01, .1), .size = c(1:10), .bag = FALSE)
nnetModel <- train(x = trainingData$x,
                   y = trainingData$y,
                   method = 'avNNet',
                   preProc = c('center', 'scale'),
                   tuneGrid = nnetGrid,
                   trControl = trainControl(method = 'cv'),
                   linout = TRUE,
                   trace = FALSE,
                   MaxNWts = 10 * (ncol(trainingData$x) + 1) + 10 + 1,
                   maxit = 500)

nnetModel
```


```{r 7.2NNetModelPredict, eval=TRUE, message=FALSE, warning=FALSE}
# Run predict() and postResample().
nnetPred <- predict(nnetModel, newdata = testData$x)
nnetPerformance <- postResample(pred = nnetPred, obs = testData$y)
nnetPerformance
```

**Model Performance Comparison**

```{r 7.2ModelPerforamnceComparison, eval=TRUE, message=FALSE, warning=FALSE}
rbind('MARS' = marsPerformance, 'SVM' = svmPerformance, 'Neural Network' = nnetPerformance, 'KNN' = knnPerformance) %>%
  kable() %>% kable_styling()
```



#### Does MARS select the informative predictors (those named X1–X5)?

```{r 7.2informativePredictors, eval=TRUE, message=FALSE, warning=FALSE}
varImp(marsModel)
```

\ 


### Exercise 7.5

Exercise 6.3 describes data for a chemical manufacturing process. Use the same data imputation, data splitting, and pre-processing steps as before and train several nonlinear regression models.

```{r 7.5, eval=TRUE, message=FALSE, warning=FALSE}
# Load the chemicalManufacturingProcess dataset that is provided by the "AppliedPredictiveModeling" package.
data(ChemicalManufacturingProcess)
```


#### Use The Same Data Imputation, Data Splitting, and Pre-Processing Steps as Before In Exercise 6.3.

**1. Impute missing values using KNN.**
```{r 7.5Impute, eval=TRUE, message=FALSE, warning=FALSE}
# Impute the missing values using KNN.
cmpImputed <- preProcess(ChemicalManufacturingProcess, 'knnImpute')
```


**2. Predict after imputation.** 
```{r 7.5Predict, eval=TRUE, message=FALSE, warning=FALSE}
# Predict after imputation.
chemicalMPData <- predict(cmpImputed, ChemicalManufacturingProcess)
```


**3. Split the data into training and test sets.**
```{r 7.5SplitData, eval=TRUE, message=FALSE, warning=FALSE}
# Split the training data using an 80% training data split.
trainingData <- createDataPartition(ChemicalManufacturingProcess$Yield, p = 0.8, list = FALSE)
xTrainData <- chemicalMPData[trainingData, ]
yTrainData <- ChemicalManufacturingProcess$Yield[trainingData]

# Split the test data.
xTestData <- chemicalMPData[-trainingData, ]
yTestData <- ChemicalManufacturingProcess$Yield[-trainingData]
```

\ 

#### Train Several Nonlinear Regression Models

##### Model 1: KNN Model

```{r 7.5KNN, eval=TRUE, message=FALSE, warning=FALSE}
set.seed(50)

# Train a KNN model.
knnModel <- train(x = xTrainData,
                  y = yTrainData,
                  method = 'knn',
                  preProc = c('center', 'scale'),
                  tuneLength = 10)
knnModel
```



```{r 7.25KNNPredict, eval=TRUE, message=FALSE, warning=FALSE}
# Run predict() and postResample().
knnPred <- predict(knnModel, newdata = xTestData)
knnPerformance <- postResample(pred = knnPred, obs = yTestData)
knnPerformance
```


**(a) Which nonlinear regression model gives the optimal resampling and test set performance?**


**(b) Which predictors are most important in the optimal nonlinear regression model? Do either the biological or process variables dominate the list? How do the top ten important predictors compare to the top ten predictors from the optimal linear model?**

```{r 7.5B, eval=TRUE, message=FALSE, warning=FALSE}

```


**(c) Explore the relationships between the top predictors and the response for the predictors that are unique to the optimal nonlinear regression model. Do these plots reveal intuition about the biological or process predictors and their relationship with yield**?

```{r 7.5C, eval=TRUE, message=FALSE, warning=FALSE}

```


\ 
