---
title: "Data 624 Homework 1"
subtitle: "Week 2 Time Series"
author: "Stephen Haslett"
date: "8/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, eval=TRUE, message=FALSE, warning=FALSE, echo=FALSE}
library(fpp3)
```


### Exercise 2.1
**Use the help function to explore what the series _gafa_stock_, _PBS_, _vic_elec_ and _pelt_ represent**.

**_gafa_stock Series Description_**
```{r exercise2.1Gafa}
help(gafa_stock)
```
_Historical stock prices from 2014-2018 for Google, Amazon, Facebook and Apple. All prices are in $USD_.

\ 

**_PBS Series Description_**
```{r exercise2.1PBS}
help(PBS)
```
_PBS is a monthly tsibble with two values_:

_Scripts: Total number of scripts_

_Cost:	Cost of the scripts in $AUD_

\ 

**_vic_elec Series Description_**
```{r exercise2.1Vic}
help(vic_elec)
```
_vic_elec is a half\-hourly tsibble with three values_:

_Demand:	Total electricity demand in MW_.

_Temperature:	Temperature of Melbourne (BOM site 086071)_.

_Holiday:	Indicator for if that day is a public holiday_.

\ 

**_pelt Series Description_**
```{r exercise2.1Pelt}
help(pelt)
```

_Hudson Bay Company trading records for Snowshoe Hare and Canadian Lynx furs from 1845 to 1935. This data contains trade records for all areas of the company_.





