---
title: "Data 624 Homework 1"
subtitle: "Week 2 Time Series"
author: "Stephen Haslett"
date: "8/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, eval=TRUE, message=FALSE, warning=FALSE, echo=FALSE}
library(fpp3)
library(DT)
```


### Exercise 2.1
**Use the help function to explore what the series _gafa_stock_, _PBS_, _vic_elec_ and _pelt_ represent**.

**_gafa_stock Series Description_**
```{r exercise2.1Gafa, eval=TRUE, message=FALSE, warning=FALSE}
help(gafa_stock)
```

_Historical stock prices from 2014-2018 for Google, Amazon, Facebook and Apple. All prices are in $USD_.


**_PBS Series Description_**
```{r exercise2.1PBS, eval=TRUE, message=FALSE, warning=FALSE}
help(PBS)
```

_PBS is a monthly tsibble with two values_:

_Scripts: Total number of scripts_

_Cost:	Cost of the scripts in $AUD_


**_vic_elec Series Description_**

```{r exercise2.1Vic, eval=TRUE, message=FALSE, warning=FALSE}
help(vic_elec)
```

_vic_elec is a half\-hourly tsibble with three values_:

_Demand:	Total electricity demand in MW_.

_Temperature:	Temperature of Melbourne (BOM site 086071)_.

_Holiday:	Indicator for if that day is a public holiday_.


**_pelt Series Description_**

```{r exercise2.1Pelt, eval=TRUE, message=FALSE, warning=FALSE}
help(pelt)
```

_Hudson Bay Company trading records for Snowshoe Hare and Canadian Lynx furs from 1845 to 1935. This data contains trade records for all areas of the company_.

\ 

**2.1 A. _Use autoplot() to plot some of the series in these data sets_**.

**_pelt Series_**

```{r exercise2.1AGafa, eval=TRUE, message=FALSE, warning=FALSE}
autoplot(pelt)
```

**_vic_elec Series_**

```{r exercise2.1AVicElec, eval=TRUE, message=FALSE, warning=FALSE}
vic_elec %>% autoplot(Demand)
```


**2.1 B. _What is the time interval of each series?_**.

* _pelt \- 1 Year_

* _vic_elec - 30 Minutes_ 

\  
\ 

### Exercise 2.2
**Use filter() to find what days corresponded to the peak closing price for each of the four stocks in gafa_stock**.

```{r exercise2.2Filter, eval=TRUE, message=FALSE, warning=FALSE}
gafa_stock %>%
  group_by(Symbol) %>%
  filter(Close == max(Close))
```

\  
\ 

### Exercise 2.3
**Download the file tute1.csv from the book website, open it in Excel (or some other spreadsheet application), and review its contents. You should find four columns of information. Columns B through D each contain a quarterly series, labelled Sales, AdBudget and GDP. Sales contains the quarterly sales for a small company over the period 1981-2005. AdBudget is the advertising budget and GDP is the gross domestic product. All series have been adjusted for inflation**.

**2.3 A. _You can read the data into R with the following script_**:

```{r exercise2.3A, eval=TRUE, message=FALSE, warning=FALSE}
tute1 <- readr::read_csv("tute1.csv")
# For readability purposes, use Data Tables to display the data rather than View().
datatable(tute1)
```

**2.3 B. _Convert the data to time series_**

```{r exercise2.3B, eval=TRUE, message=FALSE, warning=FALSE}
mytimeseries <- tute1 %>%
  mutate(Quarter = yearmonth(Quarter)) %>%
  as_tsibble(index = Quarter)
```


**2.3 C. _Construct time series plots of each of the three series_**

```{r exercise2.3C, eval=TRUE, message=FALSE, warning=FALSE}
mytimeseries %>%
  pivot_longer(-Quarter) %>%
  ggplot(aes(x = Quarter, y = value, colour = name)) +
  geom_line() +
  facet_grid(name ~ ., scales = "free_y")
```

**Check what happens when you don’t include facet_grid()**.

```{r exercise2.3FacetGrid, eval=TRUE, message=FALSE, warning=FALSE}
mytimeseries %>%
  pivot_longer(-Quarter) %>%
  ggplot(aes(x = Quarter, y = value, colour = name)) +
  geom_line()
```

**Answer**:
When _facet_grid()_ is removed, The results are displayed in one graph rather than 3 seperate graphs.

\ 
\ 

### Exercise 2.4
**The USgas package contains data on the demand for natural gas in the US**.

**2.4 A. _Install the USgas package_**.

```{r exercise2.4A, eval=TRUE, message=FALSE, warning=FALSE}
library(USgas)
```


**2.4 B. _Create a tsibble from us_total with year as the index and state as the key_**.

```{r exercise2.4B, eval=TRUE, message=FALSE, warning=FALSE}
gas_tsibble <- us_total %>% as_tsibble(key = state, index = year)
gas_tsibble
```

**2.4 C. _Plot the annual natural gas consumption by state for the New England area (comprising the states of Maine, Vermont, New Hampshire, Massachusetts, Connecticut and Rhode Island)_**.

```{r exercise2.4C, eval=TRUE, message=FALSE, warning=FALSE}
# Disable scientific numbers for readability purposes.
options(scipen = 999)

# Select New England states from the dataset.
annual_gas_consumption <- gas_tsibble %>%
  filter(state == c("Maine", "Vermont",
                    "New Hampshire", "Massachusetts",
                    "Connecticut", "Rhode Island"))

# Create the plot.
autoplot(annual_gas_consumption, y) +
  labs(title = "New England Annual Natural Gas Consumption by State",
       y = "Gas Consumption (Million Cubic Feet)",
       x = "Year")
  
```




