---
title: "Data 624 Homework 3"
subtitle: "Week 4 Forecasting"
author: "Stephen Haslett"
date: "9/14/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(fpp3)
library(DT)
library(feasts)
library(seasonal)
library(tsibble)
library(forecast)
```

### Exercise 5.1
**Produce forecasts for the following series using whichever of _NAIVE(y)_, _SNAIVE(y)_ or _RW(y ~ drift())_ is more appropriate in each case:**.

- **Australian Population (_global_economy_)**
- **Bricks (_aus_production_)**
- **NSW Lambs (_aus_livestock_)**
- **Household wealth (_hh_budget_)**.
-**Australian takeaway food turnover (_aus_retail_)**.


**_A. Australian Population (global_economy)_**.


```{r eval=TRUE, message=FALSE, warning=FALSE}

```


**_B. Bricks (aus_production)_**.


```{r eval=TRUE, message=FALSE, warning=FALSE}
```

**_C. NSW Lambs (aus_livestock)_**.


```{r eval=TRUE, message=FALSE, warning=FALSE}
```

**_D. Household wealth (hh_budget)_**.


```{r eval=TRUE, message=FALSE, warning=FALSE}
```

**_E. Australian takeaway food turnover (aus_retail)_**.


```{r eval=TRUE, message=FALSE, warning=FALSE}
```


\ 

### Exercise 5.2
**Use the Facebook stock price (_data set gafa_stock_) to do the following:**.

**a. Produce a time plot of the series**.

**b. Produce forecasts using the drift method and plot them**.

**c. Show that the forecasts are identical to extending the line drawn between the first and last observations**.

**d. Try using some of the other benchmark functions to forecast the same data set. Which do you think is best? Why**?


**_A. Produce a time plot of the series_**.

```{r 5.2A, eval=TRUE, message=FALSE, warning=FALSE}
facebook_stock <- gafa_stock %>% 
  filter(Symbol == 'FB')

facebook_stock %>% autoplot(Close) +
  labs(title = 'Facebook Stock Time Plot',
       y = 'Closing Value',
       x = 'Date')
```


**_B. Produce forecasts using the drift method and plot them**._**
```{r 5.2B, eval=TRUE, message=FALSE, warning=FALSE}
```


**_C._**
```{r eval=TRUE, message=FALSE, warning=FALSE}

```


**_D._**
```{r eval=TRUE, message=FALSE, warning=FALSE}
```

\ 

### Exercise 5.3
**Apply a seasonal naïve method to the quarterly Australian beer production data from 1992. Check if the residuals look like white noise, and plot the forecasts. The following code will help**.

_TEXT_.

```{r eval=TRUE, message=FALSE, warning=FALSE}
```


```{r eval=TRUE, message=FALSE, warning=FALSE}
```



\ 

### Exercise 5.4
**Repeat the previous exercise using the Australian Exports series from global_economy and the Bricks series from aus_production. Use whichever of NAIVE() or SNAIVE() is more appropriate in each case**.

_TEXT_.

```{r eval=TRUE, message=FALSE, warning=FALSE}

```

```{r eval=TRUE, message=FALSE, warning=FALSE}

```


```{r eval=TRUE, message=FALSE, warning=FALSE, echo=FALSE}

```

\ 

### Exercise 5.7
**For your retail time series (from Exercise 8 in Section 2.10):**.

**a. Create a training dataset consisting of observations before 2011 using**:

```{r eval=TRUE, message=FALSE, warning=FALSE, echo=TRUE}
#myseries_train <- myseries %>%
#  filter(year(Month) < 2011)
```

**b. Check that your data have been split appropriately by producing the following plot**.

```{r eval=TRUE, message=FALSE, warning=FALSE, echo=TRUE}
##autoplot(myseries, Turnover) +
  ##autolayer(myseries_train, Turnover, colour = "red")
```

**c. Fit a seasonal naïve model using SNAIVE() applied to your training data (myseries_train)**.

```{r eval=TRUE, message=FALSE, warning=FALSE, echo=TRUE}
##fit <- myseries_train %>%
  ##model(SNAIVE())
```

**d. Check the residuals**.

```{r eval=TRUE, message=FALSE, warning=FALSE, echo=TRUE}
##fit %>% gg_tsresiduals()
```
  **Do the residuals appear to be uncorrelated and normally distributed**?

**e. Produce forecasts for the test data**

```{r eval=TRUE, message=FALSE, warning=FALSE, echo=TRUE}
##fc <- fit %>%
  ##forecast(new_data = anti_join(myseries, myseries_train))
##fc %>% autoplot(myseries)

```

**f. Compare the accuracy of your forecasts against the actual values**.

```{r eval=TRUE, message=FALSE, warning=FALSE, echo=TRUE}
##fit %>% accuracy()
##fc %>% accuracy(myseries)
```

**g. How sensitive are the accuracy measures to the amount of training data used**?


**_A._**

```{r eval=TRUE, message=FALSE, warning=FALSE}

```


```{r eval=TRUE, message=FALSE, warning=FALSE}

```


```{r eval=TRUE, message=FALSE, warning=FALSE, echo=FALSE}

```

\ 


